The Analogue Method for Precipitation Forecasting: Finding Better Analogue Situations at a Sub-Daily Time Step

### 2016-05-23 Moving temporal window - HESS/template.tex
Changing mathematical variable names.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogues. The constraint being the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogues at a different hour, a sliding time window (STW) has been introduced on a reduced archive of hourly precipitation totals.

The STW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criteria on the whole distribution of analogue dates. The improvement was found to grow with the analogue ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific.

@@

The impact of the STW on the prediction performance has been assessed by means of a sub-daily precipitation series transformed into moving 24~h-totals at a 6-hourly time step. This resulted in an improvement of the prediction skills, which were even larger after recalibrating the AM parameters.

However, attempts to reconstruct longer precipitation series of running 24~h-totals by means of simple methods failed. It emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalyses through regional modeling. Then, the use of a STW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analogue situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the S1 criteria \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954}, which is a comparison of gradients, over a certain spatial window and at certain hours:

\begin{equation}
\label{eq:S1}
S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points of the gridded data describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values, the more similar the pressure fields.

@@

Then, $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day, $N_{1}$ being a parameter to calibrate. Then, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day. Following the nomenclature proposed by \citet{Horton2016}, this method will be named 2Z.

The second reference method adds a subsequent level of analogy on moisture variables, compared by means of the root-mean-square error (RMSE) (method 2Z-2MI, Table \ref{table:method_2Z-2MI}). This second predictor is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850) \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When a second level of analogy is added, a higher number of $N_{1}$ analogues is kept on the first level.

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}. AtmoSwing \citep{Horton2016}, which implements this technique, was used in order to calibrate the parameters in the present study. The score that is most often considered to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{n} \sum_{i=1}^{n} \left(  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right) 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value 1 otherwise. \citet{Bontron2004} proposed a decomposition of this score interpreted into a sharpness and an accuracy terms.

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution as the reference. The CRPSS (\textit{Continuous Ranked Probability Skill Score}) is thus defined as following:

\begin{equation}
\label{eq:CRPSS}
CRPSS = \frac{CRPS-CRPS_{r}}{CRPS_{p}-CRPS_{r}} = 1-\frac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the CRPS value for the reference (climatological distribution) and $CRPS_{p}$ would be the one for a perfect forecast (which implies $CRPS_{p}~=~0$). A better forecast is characterized by an increase in CRPSS.

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analogue situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a sliding time window (STW) was introduced to allow searching for candidates at different hours of the day.

@@

The changes due to the introduction of a STW was assessed first in regards to the analogy criteria improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, attempts were made to reconstruct longer 24~h-total precipitations series (Sect. \ref{sec:attempts}).

@@

\section{The sliding time window (STW) approach}
\label{sec:method}

@@

As sub-daily precipitation time series are usually available on a shorter period than traditional daily time steps, the first assessment consists in assessing the loss of performance resulting from a reduction of a 47~years archive (1961 to 2008) to 25~years (1982 to 2007). This change is assessed with the original method, without STW.

@@

Figure \ref{fig:changes_S1_analogues} presents the changes in the distributions of the S1 criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analogues for the Ulrichen station on the whole calibration period, due to introduction of the STW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the STW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. An increase in the difference between a fixed window and a sliding window is identifiable, which means that the last analogues are further improved. The latter effect is due to the accumulation of improvements brought by the new analogue situations in the selection.

The improvements of the S1 criteria are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the S1 criteria according to the analogue rank for the conventional method and the STW, and (bottom) quantiles of the relative reduction. This confirms that all quantile seem similarly reduced (S1 distributions keep their shape), and that this improvement is constantly increasing from the first to the last analogue (Fig. \ref{fig:changes_S1} bottom).

The median of the S1 values reduction (Fig. \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analogue and reaches more than 10~\% for the last one. This increasing trend with the analogue rank can be explained by the accumulation of better analogues in the distribution. The minimum improvement starts from 0 and reaches 2-3\%, meaning that the criteria have been improved on most analogues for every day of the calibration period. All other stations have a similar improvement of the S1 criteria, both in terms of distribution shape and amplitude.

@@

It can be assumed that the atmospheric conditions with a low dynamism, such as the frequent anticyclonic situations, will not be radically improved by the introduction of the STW. Conversely, dynamic situations, such as weather disturbances, have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. We can therefore expect to improve more significantly these situations with a higher dynamism when introducing a STW, as better matches to the target situation may be found.

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is. The results of this analysis are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations and confirms our intuition.

@@

Atmospheric dynamics varies greatly from one season to another, which reflects on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. It therefore makes sense to verify the effect of the STW separately per season.

A seasonal effect can be observed on the reduction of the S1 criteria due to the STW (Fig. \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer. One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogues are essentially found for the same hours or time windows.

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Fig. \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the STW approach, is well balanced between the 4 options. This means a change of 75\% of the analogues selection compared to the conventional approach, which improves the circulation analogy.

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogues, but their dates have changed. In contrast to earlier, both a reduction or an increase of the RMSE analogy criterion values are possible. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

This result of a globally null improvement of the RMSE values does not mean that the 2Z-2MI method cannot be improved by the STW. It means that after the selection of the analogues situations in terms of the synoptic circulation, the new candidate dates do not allow to find better analogues in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which can potentially improve the performance scores on the predictand.

@@

A systematic improvement of the S1 values was previously observed. However, finding better analogue situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the STW introduction, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

The new performance scores are provided in Table \ref{table:CRPSS_STW}, along with the differences regarding the conventional method with the same archive length. The differences ranges from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method. The introduction of the STW brings an improvement of the performance that is not very large, but that is nevertheless significant. Moreover, it requires no additional predictor. No relationship was found between the improvement of the score and the reduction of the S1 criteria, neither with the season.

@@

The S1 criteria was previously found to be improved to a greater extent for the most dynamic situations related with higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores will now be assessed regarding precipitation thresholds. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior.

@@

The previous assessment of the performance improvement was established with the original parameters optimized with a fixed time window. However, one can assume that the introduction of the STW may change the optimum of some parameters. The calibration has then been reprocessed.

After recalibrating, some changes in the optimal parameters can be observed for both 2Z (Table \ref{table:params_2Z_new}) and 2Z-2MI methods (Table \ref{table:params_2Z-2MI_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease. More significantly, the optimal number of analogues increases after introducing the STW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogues of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station. It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogues. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the S1 criteria grows along with the rank of the analogue, which shows an accumulation of better analogue situations in the distributions. It seems thus profitable to widen the selection of analogues in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogues is thus globally increased.

@@

The values of the CRPSS scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration. With the introduction of the STW, the performance loss related to the reduction of the archive length is compensated.

@@

The CRPS score can be decomposed into two components, namely sharpness and accuracy \citep[see][]{Bontron2004}. The impact of the STW has been analyzed on these components and the results can be found in Table \ref{table:CRPSS_decomposition_no_recalibration} for the original parameters and in Table \ref{table:CRPSS_decomposition_recalibration} for the re-calibrated methods. The changes are expressed relative to the total CRPS. Both components do not have the same value ranges: accuracy is in our case almost twice the sharpness values. Since the CRPS is considered here and not the CRPSS, improved prediction capacity results in a lower score. A decrease of the score is thus desirable.

It appears that the sharpness is a bit inferior with the STW in favor of accuracy (Table \ref{table:CRPSS_decomposition_no_recalibration}). When using the recalibrated parameters, the components are a bit more balanced, but still the accuracy prevails on the sharpness (Table \ref{table:CRPSS_decomposition_recalibration}). Fig. \ref{fig:CRPSS_decomposition_precip_thresholds} illustrates the changes (relative to the total CRPS) in sharpness and accuracy for different precipitation thresholds at the Ulrichen station. Be it for the 2Z or 2Z-2MI method, the original or the recalibrated parameters, the changes are similar: the improvements concern the accuracy over the sharpness, and this to a greater extent for significant precipitation events. The STW does not allow to improve the sharpness, but the improvement in accuracy is significant, and this especially for the days with heavy precipitation. In terms of the predicted precipitation distribution, it means that the median of the prediction is closer to the observed values, while the distribution is not more condensed than previously.

@@

The improvement provided by the STW is interesting, mainly for heavy precipitation events, and thus it would be profitable to be able to apply it to the complete archive. Unfortunately, there is no long precipitation time series available with a sub-daily time step. In order to reconstruct a longer archive of moving 24~h-totals, different disaggregation approaches of the daily time series were assessed and are presented in the following sections.

@@

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}). When using this reconstructed time series on the shorter period, the method performance was degraded (Table \ref{table:disaggregation_proportional} to compare to Table \ref{table:CRPSS_STW}) and was even below the conventional method without STW (Table \ref{table:loss_reduction}). The benefit of a better selection of the analogue situations is lost due to a precipitation archive of poor quality.

@@

As demonstrated in the previous section, the performance improvement brought by the STW is lost when using a simplistic precipitation time series. In order to get closer to the chronology of the actual precipitation, the reconstruction procedure must rely on some informative variable. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution.

@@

Table \ref{table:proxy_CRPSS} presents the CRPSS scores obtained by the reconstituted series using the proxy of precipitable water at the optimal point (Table \ref{table:proxy_correlations}). A slight improvement can be obtained compared to the proportional distribution method, but it is still relatively small, and most of the benefit of the STW is again lost.

These attempts to transpose the STW on the total archive highlight the importance of the actual rainfall chronology. The STW is profitable, provided that the precipitation series are close to the observed one. Without precipitation series with an accurate sub-daily chronology, the introduction of a STW remains fruitless.

@@

As \citet{Finet2008} had previously shown, the introduction of a STW allows finding better analogue situations in terms of the atmospheric circulation. It has been here demonstrated that the improvement of the S1 criteria values is growing with the rank of the analogue. This is due to the accumulation of better analogues within the predicted distributions.

The improvement of the circulation analogy was found to be more important for heavy precipitation events, which are generally related to more dynamic atmospheric situations. This dynamism implies more specific circulation patterns that are evolving more rapidly. Thus, a STW is found to be of particular interest in this kind of situation, and benefits the prediction of extreme precipitation events.

A seasonal effect has been highlighted, as the STW is more profitable for winter months than summer. The reason is likely that the diurnal cycle has a bigger effect in summer than in winter, which results in better analogues at the same time of the day. The preference for the same hours in summer has been demonstrated, whereas 75\% of the analogue situations are selected at a different time in winter.

The impact of the STW on the prediction performance has never been assessed before, due to the shortcoming of long precipitation series at a sub-daily time step. The impact of the STW approach could be assessed on a 25 years time series with a high temporal resolution. After the introduction of the STW, the performance scores increased of the same order of magnitude as if the length of the archive doubled, in this case study.

>>>>

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogues. The constraint being the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogues at a different hour, a moving time window (MTW) has been introduced on a reduced archive of hourly precipitation totals.

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criteria on the whole distribution of analogue dates. The improvement was found to grow with the analogue ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific.

@@

The impact of the MTW on the prediction performance has been assessed by means of a sub-daily precipitation series transformed into moving 24~h-totals at a 6-hourly time step. This resulted in an improvement of the prediction skills, which were even larger after recalibrating the AM parameters.

However, attempts to reconstruct longer precipitation series of running 24~h-totals by means of simple methods failed. It emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalyses through regional modeling. Then, the use of a MTW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analogue situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the \citet[Eq.\ \ref{eq:S1}, ][]{Teweles1954} score ($S_{\text{TW}}$), which is a comparison of gradients, over a certain spatial window and at certain hours:

\begin{equation}
\label{eq:S1}
S_{\text{TW}}=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points of the gridded data describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the $S_{\text{TW}}$ values, the more similar the pressure fields.

@@

Then, $N_{1}$ dates with the lowest values of $S_{\text{TW}}$ are considered as analogues to the target day, $N_{1}$ being a parameter to calibrate. Then, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day. Following the nomenclature proposed by \citet{Horton2016}, this method will be named 2Z.

The second reference method adds a subsequent level of analogy on moisture variables, compared by means of the root-mean-square error ($E_{\text{RMS}}$) (method 2Z-2MI, Table \ref{table:method_2Z-2MI}). This second predictor is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850) \citep{Bontron2004}. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When a second level of analogy is added, a higher number of $N_{1}$ analogues is kept on the first level.

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}. AtmoSwing \citep{Horton2016}, which implements this technique, was used in order to calibrate the parameters in the present study. The score that is most often considered to assess an AM performance is the Continuous Ranked Probability Score \citep[$S_{\text{CRP}}$,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. It is defined as follows:

\begin{equation}
\label{eq:CRPS}
S_{\text{CRP}} = \int_{-\infty}^{+\infty} \left[ F(y)-\text{H}(y-y^{0})\right]^{2} \text{d}y ,  
\end{equation}
where $\text{H}(y-y^{0})$ is the Heaviside function that is null when $y-y^{0}<0$, and has the value 1 otherwise. \citet{Bontron2004} proposed a decomposition of this score interpreted into a sharpness and an accuracy terms.

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution as the reference. The Continuous Ranked Probability Skill Score ($S_{\text{SCRP}}$) is thus defined as following:

\begin{equation}
\label{eq:CRPSS}
S_{\text{SCRP}} = \frac{S_{\text{CRP}}-S_{\text{CRP}}^{r}}{S_{\text{CRP}}^{p}-S_{\text{CRP}}^{r}} = 1-\frac{S_{\text{CRP}}}{S_{\text{CRP}}^{r}}
\end{equation}
where $S_{\text{CRP}}^{r}$ is the $S_{\text{CRP}}$ value for the reference (climatological distribution) and $S_{\text{CRP}}^{p}$ would be the one for a perfect forecast (which implies $S_{\text{CRP}}^{p}~=~0$). A better forecast is characterized by an increase in $S_{\text{SCRP}}$.

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analogue situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a moving time window (MTW) was introduced to allow searching for candidates at different hours of the day.

@@

The changes due to the introduction of a MTW was assessed first in regards to the analogy criteria improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, attempts were made to reconstruct longer 24~h-total precipitations series (Sect. \ref{sec:attempts}).

@@

\section{The moving time window (MTW) approach}
\label{sec:method}

@@

As sub-daily precipitation time series are usually available on a shorter period than traditional daily time steps, the first assessment consists in assessing the loss of performance resulting from a reduction of a 47~years archive (1961 to 2008) to 25~years (1982 to 2007). This change is assessed with the original method, without MTW.

@@

Figure \ref{fig:changes_S1_analogues} presents the changes in the distributions of the $S_{\text{TW}}$ criterion for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analogues for the Ulrichen station on the whole calibration period, due to introduction of the MTW. The precipitation target remains as before, that is centered on 18~h~UTC (6~h~UTC to 6~h~UTC the next day). The shapes of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criteria are now reduced (shifted to the left), and therefore better. An increase in the difference between a fixed window and a moving window is identifiable, which means that the last analogues are further improved. The latter effect is due to the accumulation of improvements brought by the new analogue situations in the selection.

The improvements of the $S_{\text{TW}}$ criteria are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the $S_{\text{TW}}$ criteria according to the analogue rank for the conventional method and the MTW, and (bottom) quantiles of the relative reduction. This confirms that all quantile seem similarly reduced ($S_{\text{TW}}$ distributions keep their shape), and that this improvement is constantly increasing from the first to the last analogue (Fig. \ref{fig:changes_S1} bottom).

The median of the $S_{\text{TW}}$ values reduction (Fig. \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analogue and reaches more than 10~\% for the last one. This increasing trend with the analogue rank can be explained by the accumulation of better analogues in the distribution. The minimum improvement starts from 0 and reaches 2-3\%, meaning that the criteria have been improved on most analogues for every day of the calibration period. All other stations have a similar improvement of the $S_{\text{TW}}$ criteria, both in terms of distribution shape and amplitude.

@@

It can be assumed that the atmospheric conditions with a low dynamism, such as the frequent anticyclonic situations, will not be radically improved by the introduction of the MTW. Conversely, dynamic situations, such as weather disturbances, have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. We can therefore expect to improve more significantly these situations with a higher dynamism when introducing a MTW, as better matches to the target situation may be found.

The dynamism of a given atmospheric condition cannot be easily quantified. A basic assumption is considered here, which stipulates that the more a day is rainy, the more dynamic the situation is. The results of this analysis are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of $S_{\text{TW}}$ for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations and confirms our intuition.

@@

Atmospheric dynamics varies greatly from one season to another, which reflects on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. It therefore makes sense to verify the effect of the MTW separately per season.

A seasonal effect can be observed on the reduction of the $S_{\text{TW}}$ criteria due to the MTW (Fig. \ref{fig:changes_S1_seasons}). The improvements are greater for winter than summer. One hypothesis is that the diurnal effects of the summer months have an influence on the atmospheric circulation, at least in the lower layers. This effect is based on the daily cycle and good analogues are essentially found for the same hours or time windows.

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Fig. \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the MTW approach, is well balanced between the 4 options. This means a change of 75\% of the analogues selection compared to the conventional approach, which improves the circulation analogy.

@@

When considering the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidates situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogues, but their dates have changed. In contrast to earlier, both a reduction or an increase of the $E_{\text{RMS}}$ analogy criterion values are possible. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the $E_{\text{RMS}}$ values. Unlike for the first level of analogy, the relative improvement of the $E_{\text{RMS}}$ values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

This result of a globally null improvement of the $E_{\text{RMS}}$ values does not mean that the 2Z-2MI method cannot be improved by the MTW. It means that after the selection of the analogues situations in terms of the synoptic circulation, the new candidate dates do not allow to find better analogues in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which can potentially improve the performance scores on the predictand.

@@

A systematic improvement of the $S_{\text{TW}}$ values was previously observed. However, finding better analogue situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the MTW introduction, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see Sect. \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

The new performance scores are provided in Table \ref{table:CRPSS_MTW}, along with the differences regarding the conventional method with the same archive length. The differences ranges from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method. The introduction of the MTW brings an improvement of the performance that is not very large, but that is nevertheless significant. Moreover, it requires no additional predictor. No relationship was found between the improvement of the score and the reduction of the $S_{\text{TW}}$ criteria, neither with the season.

@@

The $S_{\text{TW}}$ criteria was previously found to be improved to a greater extent for the most dynamic situations related with higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores will now be assessed regarding precipitation thresholds. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior.

@@

The previous assessment of the performance improvement was established with the original parameters optimized with a fixed time window. However, one can assume that the introduction of the MTW may change the optimum of some parameters. The calibration has then been reprocessed.

After recalibrating, some changes in the optimal parameters can be observed for both 2Z (Table \ref{table:params_2Z_new}) and 2Z-2MI methods (Table \ref{table:params_2Z-2MI_new}). Among these, the east-west dimension of the spatial windows of the circulation analogy tends to decrease. More significantly, the optimal number of analogues increases after introducing the MTW, of a significant order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method. The number of analogues of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station. It seems as if the method perceive the increase in the archive length and diversity by allowing to extract more analogues. As we saw in Fig. \ref{fig:changes_S1}, the improvement of the $S_{\text{TW}}$ criteria grows along with the rank of the analogue, which shows an accumulation of better analogue situations in the distributions. It seems thus profitable to widen the selection of analogues in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogues is thus globally increased.

@@

The values of the $S_{\text{SCRP}}$ scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration. With the introduction of the MTW, the performance loss related to the reduction of the archive length is compensated.

@@

The $S_{\text{CRP}}$ score can be decomposed into two components, namely sharpness and accuracy \citep[see][]{Bontron2004}. The impact of the MTW has been analyzed on these components and the results can be found in Table \ref{table:CRPSS_decomposition_no_recalibration} for the original parameters and in Table \ref{table:CRPSS_decomposition_recalibration} for the re-calibrated methods. The changes are expressed relative to the total $S_{\text{CRP}}$. Both components do not have the same value ranges: accuracy is in our case almost twice the sharpness values. Since the $S_{\text{CRP}}$ is considered here and not the $S_{\text{SCRP}}$, improved prediction capacity results in a lower score. A decrease of the score is thus desirable.

It appears that the sharpness is a bit inferior with the MTW in favor of accuracy (Table \ref{table:CRPSS_decomposition_no_recalibration}). When using the recalibrated parameters, the components are a bit more balanced, but still the accuracy prevails on the sharpness (Table \ref{table:CRPSS_decomposition_recalibration}). Fig. \ref{fig:CRPSS_decomposition_precip_thresholds} illustrates the changes (relative to the total $S_{\text{CRP}}$) in sharpness and accuracy for different precipitation thresholds at the Ulrichen station. Be it for the 2Z or 2Z-2MI method, the original or the recalibrated parameters, the changes are similar: the improvements concern the accuracy over the sharpness, and this to a greater extent for significant precipitation events. The MTW does not allow to improve the sharpness, but the improvement in accuracy is significant, and this especially for the days with heavy precipitation. In terms of the predicted precipitation distribution, it means that the median of the prediction is closer to the observed values, while the distribution is not more condensed than previously.

@@

The improvement provided by the MTW is interesting, mainly for heavy precipitation events, and thus it would be profitable to be able to apply it to the complete archive. Unfortunately, there is no long precipitation time series available with a sub-daily time step. In order to reconstruct a longer archive of moving 24~h-totals, different disaggregation approaches of the daily time series were assessed and are presented in the following sections.

@@

A proportional distribution is certainly the simplest approach that can be performed. It consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}). When using this reconstructed time series on the shorter period, the method performance was degraded (Table \ref{table:disaggregation_proportional} to compare to Table \ref{table:CRPSS_MTW}) and was even below the conventional method without MTW (Table \ref{table:loss_reduction}). The benefit of a better selection of the analogue situations is lost due to a precipitation archive of poor quality.

@@

As demonstrated in the previous section, the performance improvement brought by the MTW is lost when using a simplistic precipitation time series. In order to get closer to the chronology of the actual precipitation, the reconstruction procedure must rely on some informative variable. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution.

@@

Table \ref{table:proxy_CRPSS} presents the $S_{\text{SCRP}}$ scores obtained by the reconstituted series using the proxy of precipitable water at the optimal point (Table \ref{table:proxy_correlations}). A slight improvement can be obtained compared to the proportional distribution method, but it is still relatively small, and most of the benefit of the MTW is again lost.

These attempts to transpose the MTW on the total archive highlight the importance of the actual rainfall chronology. The MTW is profitable, provided that the precipitation series are close to the observed one. Without precipitation series with an accurate sub-daily chronology, the introduction of a MTW remains fruitless.

@@

As \citet{Finet2008} had previously shown, the introduction of a MTW allows finding better analogue situations in terms of the atmospheric circulation. It has been here demonstrated that the improvement of the $S_{\text{TW}}$ criteria values is growing with the rank of the analogue. This is due to the accumulation of better analogues within the predicted distributions.

The improvement of the circulation analogy was found to be more important for heavy precipitation events, which are generally related to more dynamic atmospheric situations. This dynamism implies more specific circulation patterns that are evolving more rapidly. Thus, a MTW is found to be of particular interest in this kind of situation, and benefits the prediction of extreme precipitation events.

A seasonal effect has been highlighted, as the MTW is more profitable for winter months than summer. The reason is likely that the diurnal cycle has a bigger effect in summer than in winter, which results in better analogues at the same time of the day. The preference for the same hours in summer has been demonstrated, whereas 75\% of the analogue situations are selected at a different time in winter.

The impact of the MTW on the prediction performance has never been assessed before, due to the shortcoming of long precipitation series at a sub-daily time step. The impact of the MTW approach could be assessed on a 25 years time series with a high temporal resolution. After the introduction of the MTW, the performance scores increased of the same order of magnitude as if the length of the archive doubled, in this case study.
